### 2019-01-31 abstract.tex
ssr

A class of Boolean functions, 
called \textbf{nested canalyzing functions} (NCFs),
has been used to model certain biological phenomena.
Researchers have studied some properties of these functions
such as sensitivity and stability.
In this note, we examine the relationships between NCFs and symmetric 
Boolean functions. 
We identify all the Boolean functions that both  NCFs and symmetric. 
A more general form of symmetric Boolean functions,
called $r$-symmetric functions (where $r$ is the symmetry level),
have also been studied in the literature.
%% We establish lower and upper bounds on the symmetry
%% level of an NCF $f$ as functions of the number of layers
%% in a standard respresentation of $f$.  
We show that, in general,
one cannot efficiently approximate the symmetry level of
a Boolean function to within any factor, unless \textbf{P} = \cnp. 
In contrast, we show that the symmetry level of an NCF $f$
can be easily computed from a standard (and efficiently
computable) representation of $f$.

>>>>

A class of Boolean functions, 
called \textbf{nested canalyzing functions} (NCFs),
has been used to model certain biological phenomena.
Researchers have studied some properties of these functions
such as sensitivity and stability.
In this note, we examine the relationships between NCFs and symmetric 
Boolean functions. 
We identify all the Boolean functions that NCFs as well as symmetric. 
A more general form of symmetric Boolean functions,
called $r$-symmetric functions (where $r$ is the symmetry level),
has also been studied in the literature.
%% We establish lower and upper bounds on the symmetry
%% level of an NCF $f$ as functions of the number of layers
%% in a standard respresentation of $f$.  
We observe that, in general,
one cannot efficiently approximate the symmetry level of
a Boolean function to within any factor, unless \textbf{P} = \cnp. 
In contrast, we show that the symmetry level of an NCF $f$
can be easily computed from a standard (and efficiently
computable) representation of $f$.

### 2019-01-31 intro.tex
ssr

The following definition of \textbf{canalyzing} Boolean functions was introduced
in \cite{Kauffman-1969}.

\begin{definition}\label{def:canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ be a set of $n$  Boolean variables.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{canalyzing}
if there is a variable $x_i \in X$ and values $a$, $b$ $\in \{0,1\}$ such that 
$f(x_1, \ldots, x_{i-1}, a, x_{i+1}, \ldots, x_n)$ $~=~ b$, regardless of the values
assigned to the variables in $X - \{x_i\}$.
\end{definition}

@@

Another class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
was introduced later in \cite{Kauffman-etal-2003} to model the behavior of
certain biological systems.
We follow the presentation in \cite{Layne-2011} in defining NCFs.
(For a Boolean value $b$,~ the complement is denoted by $\overline{b}$.)

\begin{definition}\label{def:nested_canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ denote a set of $n$  Boolean variables.
Let $\pi$ be a permutation of $\{1, 2, \ldots, n\}$.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{nested canalyzing}
in the variable order $x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}$ with
\textbf{canalyzing values} $a_1, a_2, \ldots, a_n$ and 
\textbf{canalyzed values} $b_1, b_2, \ldots, b_n$ 
if $f$ can be expressed in the following form:
\[
f(x_1, x_2, \ldots, x_n) ~=~ 
   \begin{cases}
       \:b_1 & \mathrm{if~~} x_{\pi(1)} ~=~ a_1 \\
       \:b_2 & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~}
            x_{\pi(2)} ~=~ a_2 \\
       \:\vdots & \vdots \\
       \:b_n & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
             x_{\pi(n-1)} ~\neq~ a_{n-1} \mathrm{~~and~~} x_{\pi(n)} ~=~ a_n \\
       \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
            x_{\pi(n)} ~\neq~ a_n \\
   \end{cases}
\]
\end{definition}
%% \clearpage
For convenience, we will use a notation introduced in \cite{Stearns-etal-2018}
to represent NCFs.
For $1 \leq i \leq n$, line $i$ of this representation has the following form:

@@

\noindent 
Following the terminology in \cite{Stearns-etal-2018},
 $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
\textbf{tested} in line $i$, 
with $a_i$ and $b_i$ denoting respectively the canalyzing and 
canalyzed values in line $i$ as before,~ $1 \leq i \leq n$.
Each such line is called a \textbf{rule}.
When none of the conditions ``$x_{\pi(i)} ~=~ a_i$" 
is satisfied, we have line $n+1$ with the ``Default" rule
for which the canalyzed value is~ $\overline{b_n}$:

@@

A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
said to be \textbf{symmetric} if their values can be interchanged without
affecting the value of the function.
As a simple example, each pair of variables in the function 
$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ is symmetric.
This form of interchange symmetry partitions the set of variables into a set of
\textbf{symmetry groups}, where the members of each group are pairwise symmetric.
Function $f$ is said to be
$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
In this case, the value of $f$
depends only on how many of the variables in each symmetry group have the value 1.
We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
it is $r$-symmetric, but not $(r-1)$-symmetric.
For example, the function $f_3(x_1, x_2) = x_1 \vee \overline{x_2}$
is not symmetric since $f(1, 0) \neq f(0, 1)$; however, 
it is trivially 2-symmetric.
For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
will be referred to as the \textbf{symmetry level} of $f$.
(Thus, the symmetry level of $f$ is the smallest integer $r$ such that
$f$ is $r$-symmetric.)

@@

The term \textbf{canalization} was first used by
Conrad Waddington (a British geneticist)
in 1942 \cite{Waddington-1942}
to express the ability of a species
to produce members with very similar observable features over a period
of time even when there are significant changes in the environment
during that period.
Since that time, the term ``canalization'' has been used to describe
the stability of a biological system with changes
in external conditions.
In 1969, Kauffman \cite{Kauffman-1969} introduced a Boolean network model
to explain the stability of gene regulatory networks.
Kauffman observed that the use of canalyzing Boolean
functions in the model captured many observed properties of
gene regulatory networks.
The subclass of NCFs was
introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
to facilitate a rigorous analysis of the Boolean network model
for gene regulatory networks.
It is known that the class of NCFs coincides with that
of unate cascade functions \cite{}
Many researchers have studied mathematical properties of NCFs and
pointed out the importance of NCFs in modeling biological
phenomena (e.g., \cite{Layne-2011,
Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
Additionally, properties of NCFs such as sensitivity and 
stability have also been studied
by many researchers 
\cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
Li-etal-2011,Li-etal-2013,Klotz-etal-2013,
Stearns-etal-2018}. 
To our knowledge, the relationship between NCFs and symmetric
Boolean functions has not been studied in the literature.

>>>>

The class of \textbf{canalyzing} Boolean functions, introduced 
in \cite{Kauffman-1969}, is defined as follows.

\begin{definition}\label{def:canalyzing}
Given a set $X = $ $\{x_1, x_2, \ldots, x_n\}$ of $n$  Boolean variables,
a Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{canalyzing}
if there is a variable $x_i \in X$ and values $a$, $b$ $\in \{0,1\}$ such that
\[
f(x_1, \ldots, x_{i-1}, a, x_{i+1}, \ldots, x_n) ~=~ b, 
\]
regardless of the values assigned to the variables in $X - \{x_i\}$.
\end{definition}

@@

Another class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
was introduced later in \cite{Kauffman-etal-2003} to carry out a detailed
analysis of the behavior of certain biological systems.
We follow the presentation in \cite{Layne-2011} in defining NCFs.
(For a Boolean value $b$,~ the complement is denoted by $\overline{b}$.)

\begin{definition}\label{def:nested_canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ denote a set of $n$  Boolean variables.
Let $\pi$ be a permutation of $\{1, 2, \ldots, n\}$.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{nested canalyzing}
in the variable order $x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}$ with
\textbf{canalyzing values} $a_1, a_2, \ldots, a_n$ and 
\textbf{canalyzed values} $b_1, b_2, \ldots, b_n$ 
if $f$ can be expressed in the following form:
\[
f(x_1, x_2, \ldots, x_n) ~=~ 
   \begin{cases}
       \:b_1 & \mathrm{if~~} x_{\pi(1)} ~=~ a_1 \\
       \:b_2 & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~}
            x_{\pi(2)} ~=~ a_2 \\
       \:\vdots & \vdots \\
       \:b_n & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
             x_{\pi(n-1)} ~\neq~ a_{n-1} \mathrm{~~and~~} x_{\pi(n)} ~=~ a_n \\
       \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
            x_{\pi(n)} ~\neq~ a_n \\
   \end{cases}
\]
\end{definition}
%% \clearpage
For convenience, we will use a notation introduced in \cite{Stearns-etal-2018}
to represent NCFs.
For $1 \leq i \leq n$, line $i$ of this representation has the form

@@

\noindent 
with $x_{\pi(i)}$ being the \textbf{canalyzing variable} that is
\textbf{tested} in line $i$, 
and $a_i$ and $b_i$ being respectively the canalyzing and 
canalyzed values in line $i$,~ $1 \leq i \leq n$.
Each such line is called a \textbf{rule}.
When none of the conditions ``$x_{\pi(i)} ~=~ a_i$" 
is satisfied, we have line $n+1$ with the ``Default" rule
for which the canalyzed value is~ $\overline{b_n}$:

@@

A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
said to be \textbf{symmetric} if their values can be interchanged without
affecting the value of the function.
As a simple example, each pair of variables in the function 
$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ is symmetric.
This form of interchange symmetry partitions the set of variables into a set of
\textbf{symmetry groups}, where the members of each group are pairwise symmetric.
Function $f$ is said to be
$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
In this case, the value of $f$
depends only on how many of the variables in each symmetry group have the value 1.
We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
it is $r$-symmetric, but not $(r-1)$-symmetric.
For example, the function $f_3(x_1, x_2) = x_1 \vee\, \overline{x_2}$
is not 1-symmetric since $f_3(1, 0) \neq f_3(0, 1)$; however, 
it is trivially 2-symmetric.
For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
will be referred to as the \textbf{symmetry level} of $f$.
Thus, the symmetry level of $f$ is the smallest integer $r$ such that
$f$ is $r$-symmetric.

@@

The term \textbf{canalization}, first used by
Waddington \cite{Waddington-1942}, is generally used to describe
the stability of a biological system with changes
in external conditions.
In 1969, Kauffman \cite{Kauffman-1969} introduced canalyzing Boolean functions
to explain the stability of gene regulatory networks.
The subclass of NCFs was
introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
to facilitate a rigorous analysis of the Boolean network model
for gene regulatory networks.
It is known that the class of NCFs coincides with that
of unate cascade Boolean functions \cite{Jarrah-etal-2007}.
In the literature on computational learning theory,
generalized NCFs are referred to as $1$-\textbf{decision lists} \cite{KV-1994}.
Many researchers have pointed out the importance of NCFs 
in modeling biological phenomena 
(e.g., \cite{Layne-2011,
Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
Additionally, properties of NCFs such as sensitivity and 
stability have also been studied
by many researchers 
\cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
Li-etal-2011,Li-etal-2013,Klotz-etal-2013,
Stearns-etal-2018}. 
To our knowledge, the relationship between NCFs and symmetric
Boolean functions has not been studied in the literature.

### 2019-01-31 preliminaries.tex
ssr

We begin with a two-part observation from \cite{Stearns-etal-2018}.

@@

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_4:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_5:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_6:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & Default:~ $1$ \\
\end{tabular}

\medskip
\noindent
This function has 4 layers, with 
Layer~1 consisting of the lines with $x_1$ and $x_2$, 
Layer~2 consisting of the lines with $x_3$ and $x_4$, 
Layer~3 consisting of the line with $x_5$ and 
Layer~4 consisting of the line with $x_6$. \qed

@@

\begin{definition}
\label{def:normalized}
Let $f$ be an NCF with $n \geq 2$ variables specified using
the simplified representation with 
lexicographic ordering given by  
$\langle x_{\pi(1)}, x_{\pi(2)}, x_{\pi(n)}\rangle$.
We say that the representation is {\bf normalized} if
both of the following conditions hold.
\begin{enumerate}
  \item Lines $n-1$ and $n$ have the same canalyzed output ~and
  \item each pair of consecutive lines
        with the same canalyzed output occur in the lexicographic order 
        of their canalyzing variables; that is, if two consecutive lines are \\
%\medskip
\noindent
\hspace*{1.1in} $x_i$:~ $a_i \longrightarrow b$ \\
\hspace*{1.1in} $x_j$:~ $a_j \longrightarrow b$

@@

\subsection{Complexity of Estimating the Symmetry Level of a i
Boolean Function}
\label{sse:symmetry_level_hardness}

Recall that the symmetry level of a Boolean function $f$ is the smallest
integer $r$ such that the following condition holds:
the  inputs to $f$ can be partitioned into $r$ subsets 
(symmetry groups) so that the value of $f$ depends only on 
how many of the inputs in each group have the value 1.
The following result shows that given a Boolean function $f$
in the form of a Boolean expression, it is \cnp-hard to 
approximate\footnote{An algorithm approximates the symmetry
level of a Boolean function $f$ within the factor $\rho$ if it finds a
partition of the inputs to $f$ into at most $\rho \times r^*$ symmetry
groups, where $r^*$ is the symmetry level of $f$.}
the symmetry level of $f$ to within any factor $\rho \geq 1$.

@@

\begin{theorem}\label{thm:approx_sym_level_hard}
Unless \textbf{P} = \cnp,
for any $\rho \geq 1$, there is no polynomial time 
algorithm for approximating the 
symmetry level of a Boolean function $f$ to within the factor $\rho$.
This result holds even when $f$ is a CNF expression.
\end{theorem}

\noindent
\textbf{Proof:}~ Suppose there is an efficient algorithm \cala{} 
that provides a $\rho$ approximation for the symmetry level
of a Boolean function for some $\rho \geq 1$.
We will show that \cala{} can be used to efficiently solve
the CNF Satisfiability problem (SAT) which is known 
to be \cnp-hard \cite{GJ-1979}.

@@

\noindent
\textbf{Proof of Claim 1:}~ If $f$ is not satisfiable, $f'$ is also
not satisfiable; that is, for all inputs, the value of $f'$ is 0.
Thus, all the variables in $X \cup Y \cup Z$ can be put in one
symmetry group; regardless of the number of variables 
which are 1 in the group, the value of $f'$ is always 0. 
\hfill$\Box$

@@

\noindent
\textbf{Proof of Claim 2:}~ Suppose $f$ is satisfiable.
We argue that for $i \neq j$, variables $y_i$ and $y_j$
cannot be in the same symmetry group.
To see this, consider the following assignment $\alpha$ to
the variables of $f'$.
\begin{description}
\item{(i)} For the variables in $X$, choose an assignment that
satisfies $f$.
\item{(ii)} Let $y_i = 0$, $z_i = 1$, $y_j = 1$ and $z_j = 0$.
\item{(iii)} For $1 \leq p \leq \rho+1$, $p \neq i$ and $p \neq j$,
let $y_p = 1$ and $z_p$ = 0.
\end{description}
Since this assignment $\alpha$ sets the clause $(y_i \vee \overline{z_i})$ to 0,
the value of $f'$ under the assignment $\alpha$ is 0.
Now, consider the assignment $\alpha'$ that is obtained by exchanging
the values of $y_i$ and $y_j$ in $\alpha$ without changing the
values assigned to the other variables. 
It can be seen that under assignment $\alpha'$, the value of $f'$ is 1.
In both $\alpha$ and $\alpha'$, the number of 1-valued variables in 
the set $\{y_i, y_j\}$ is 1; however, the two assignments lead to 
different values for $f'$.
Therefore, if $f$ is not satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
be in the same symmetry group.
Since there are $\rho+1$ variables in $Y$, the number of symmetry groups 
for $f'$ must be at least $\rho+1$. \hfill$\Box$

@@

Now, suppose we give $f'$ as the input to Algorithm \cala.
If $f$ is \emph{not} satisfiable, then from Claim 1, 
the symmetry level of $f'$ is 1. 
Since \cala{} is a $\rho$-approximation algorithm,
it should produce at most $\rho$ symmetry groups.
On the other hand, if $f$ is satisfiable, by Claim 2, the symmetry
level of $f'$ is at least $\rho+1$ and \cala{} should produce at least
$\rho+1$ groups.
Thus, by examining the output of \cala, we can determine whether or not $f$
is satisfiable.
Since \cala{} runs in polynomial time, we have an efficient 
algorithm for SAT, contradicting the assumption that \textbf{P} $\neq$ \cnp.
\QED

@@

In contrast to the above result,
we will show that the symmetry level of an NCF (or that of a
generalized NCF) can be computed efficiently.

>>>>

We begin with an observation from \cite{Stearns-etal-2018}.

@@

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_4:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_5:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_6:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & Default:~ $1$ \\
\end{tabular}

\medskip
\noindent
This function has 3 layers, with 
Layer~1 consisting of the lines with $x_1$, $x_2$ and $x_3$, 
Layer~2 consisting of the lines with $x_4$ and $x_5$, and
Layer~3 consisting of the line with $x_6$.  \qed

@@

\begin{definition}
\label{def:normalized}
Let $f$ be an NCF with $n \geq 2$ variables specified using
the simplified representation with 
lexicographic ordering given by  
$\langle x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}\rangle$.
We say that the representation is {\bf normalized} if
both of the following conditions hold.
\begin{enumerate}
  \item Lines $n-1$ and $n$ have the same canalyzed output ~and
  \item each pair of consecutive lines
        with the same canalyzed output occur in the lexicographic order 
        of their canalyzing variables; that is, if two consecutive lines are \\
%\medskip
\noindent
\hspace*{1.1in} $x_i$:~ $a_i \longrightarrow b$ \\
\hspace*{1.1in} $x_j$:~ $a_j \longrightarrow b$

@@

\subsection{Complexity of Estimating the Symmetry Level of a 
Boolean Function}
\label{sse:symmetry_level_hardness}

Recall that the symmetry level of a Boolean function $f$ is the smallest
integer $r$ such that the following condition holds:
the  inputs to $f$ can be partitioned into $r$ subsets 
(symmetry groups) so that the value of $f$ depends only on 
how many of the inputs in each group have the value 1.
We now show that given a Boolean function $f$
in the form of a Boolean expression, it is \cnp-hard to 
approximate\footnote{An algorithm approximates the symmetry
level of a Boolean function $f$ within the factor $\rho$ if it finds a
partition of the inputs to $f$ into at most $\rho \times r^*$ symmetry
groups, where $r^*$ is the symmetry level of $f$.}
the symmetry level of $f$ to within any factor $\rho \geq 1$.

@@

\begin{theorem}\label{thm:approx_sym_level_hard}
Unless \textbf{P} = \cnp,
for any $\rho \geq 1$, there is no polynomial time 
algorithm for approximating the 
symmetry level of a Boolean function $f$ 
specified as a Boolean expression to within the factor $\rho$.
This result holds even when $f$ is a CNF expression.
\end{theorem}

\noindent
\textbf{Proof:}~ For some $\rho \geq 1$, suppose there is an efficient algorithm \cala{} 
that provides a $\rho$ approximation for the symmetry level
of any Boolean function specified as a Boolean expression.
We will show that \cala{} can be used to efficiently solve
the CNF Satisfiability problem (SAT) which is known 
to be \cnp-hard \cite{GJ-1979}.

@@

\noindent
\textbf{Proof of Claim 1:}~ If $f$ is not satisfiable, $f'$ is also
not satisfiable; that is, for all inputs, the value of $f'$ is 0.
Thus, all the variables in $X \cup Y \cup Z$ can be included in one
symmetry group.
The value of $f'$ is 0 regardless of the number of variables 
which are 1 in the group. 
\hfill$\Box$

@@

\noindent
\textbf{Proof of Claim 2:}~ Suppose $f$ is satisfiable.
We argue that if $i \neq j$, variables $y_i$ and $y_j$
cannot be in the same symmetry group for the function $f'$.
To see this, consider any two variables $y_i$ and $y_j$ with $i \neq j$,
and construct an assignment $\alpha$ to
the variables of $f'$ in the following manner.
\begin{description}
\item{(i)} For the variables in $X$, choose an assignment that
satisfies $f$.
\item{(ii)} Let $y_i = 0$, $z_i = 1$, $y_j = 1$ and $z_j = 0$.
\item{(iii)} For $1 \leq p \leq \rho+1$, $p \neq i$ and $p \neq j$,
let $y_p = 1$ and $z_p$ = 0.
\end{description}
Since this assignment $\alpha$ sets the clause $(y_i \vee \overline{z_i})$ to 0,
the value of $f'$ under the assignment $\alpha$ is 0.
Now, consider the assignment $\alpha'$ that is obtained by interchanging
the values of $y_i$ and $y_j$ in $\alpha$ without changing the
values assigned to the other variables. 
It can be seen that under assignment $\alpha'$, the value of $f'$ is 1.
In both $\alpha$ and $\alpha'$, the number of 1-valued variables in 
the set $\{y_i, y_j\}$ is 1; however, the two assignments lead to 
different values for $f'$.
Therefore, if $f$ is not satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
be in the same symmetry group.
Since there are $\rho+1$ variables in $Y$, the number of symmetry groups 
for $f'$ must be at least $\rho+1$. \hfill$\Box$

@@

Now, suppose we execute Algorithm \cala{} on the function $f'$.
If $f$ is \emph{not} satisfiable, then from Claim 1, 
the symmetry level of $f'$ is 1. 
Since \cala{} is a $\rho$-approximation algorithm,
it should produce at most $\rho$ symmetry groups.
On the other hand, if $f$ is satisfiable, by Claim 2, the symmetry
level of $f'$ is at least $\rho+1$; so, \cala{} will produce at least
$\rho+1$ groups.
Thus, by examining the output of \cala, we can determine whether or not $f$
is satisfiable.
Since \cala{} runs in polynomial time, we have an efficient 
algorithm for SAT, contradicting the assumption that \textbf{P} $\neq$ \cnp.
\QED

@@

In contrast to the above result,
we will show in the next section that the symmetry level 
of an NCF (or that of a
generalized NCF) can be computed efficiently.
